---
title: "Course Work - Cyber Security: Safety At Home, Online, and in Life"
author: "Varsha Venkataraman - 240150459 "
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup,message=FALSE, warning=FALSE, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE)  # Disable caching globally
knitr::opts_knit$set(root.dir = normalizePath(".."))
```
```{r,message=FALSE, warning=FALSE, echo = FALSE}
library(ProjectTemplate)
load.project()
filtered_data <- combined_data %>%
    filter(!is.na(employment_area))
ggplot(filtered_data, aes(x = reorder(employment_area, employment_area, function(x) -length(x)))) +
    geom_bar(fill = "steelblue") +
    labs(
        title = "Enrollment in Cyber Security Course by Employment Area",
        x = "Employment Area",
        y = "Number of Enrollments"
    ) +
    theme_minimal() +
    theme(
        axis.text.x = element_text(angle = 45, hjust = 1, size = 8),  # Adjust font size and angle
        plot.title = element_text(hjust = 0.5)
    ) +
    scale_y_continuous(expand = expansion(mult = c(0, 0.05)))
```

The above bar plot represents the 'Employment Area' of the learners who enrolled in the cyber security course in the university. The y axis has the different type of 'Employment Area' of the learners. The x axis has the number of enrollments in the cyber security course which has the range from 0 to 1500. 

From the visualization we can conclude that the learners from IT industry has the maximum enrollment and the learners from the recruitment industry has the least enrollment. It is very obvious that the IT field is related to the course, but its surprising that the Finance area is standing seventh in the column. 

The cyber security has important role in the finance and banking area as many fraudulent acts occurs in this area. 

```{r,message=FALSE, warning=FALSE, echo = FALSE}
total_learners <- nrow(combined_data)
total_learners_df <- data.frame(
  Metric = "Total Learners",
  Value = total_learners
)

# Unique Learners
unique_learners <- n_distinct(combined_data$learner_id)
unique_learners_df <- data.frame(
  Metric = "Unique Learners",
  Value = unique_learners
)

# Combine the tables side by side using `cbind`
combined_table <- cbind(total_learners_df, unique_learners_df)

# Display the result in a side-by-side table format using kable
kable(combined_table, caption = "Learner Metrics") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```
```{r,message=FALSE, warning=FALSE, echo = FALSE}
unique_learners <- n_distinct(combined_data$learner_id)

# Create a data frame for displaying the count of unique learners
unique_learners_df <- data.frame(
  Metric = "Unique Learners",
  Value = unique_learners
)

# Display the result in a table format using kable
kable(unique_learners_df, caption = "Number of Unique Learners")
```

With respect to the above Table 1, the total number of learners who are enrolled in the cyber security course is 63,248 of them. Looking at the pie chart below which has the percentage value for the learners per employment area. We can conclude that we have the employment area information only for 9% of the learners only.

```{r,message=FALSE, warning=FALSE, echo = FALSE,fig.height=5,fig.width=5,fig.align='center'}
colors <- colorRampPalette(brewer.pal(9, "Set1"))(22)
summary_data <- combined_data %>%
    count(employment_area) %>%
    rename(Count = n) %>%
    add_row(employment_area = "Missing", Count = sum(is.na(combined_data$employment_area))) %>%
    add_row(employment_area = "Total", Count = nrow(combined_data)) %>%
    filter(employment_area != "Total")  # Exclude the total row for pie chart clarity

# Assign custom labels for the segments
summary_data$label <- ifelse(summary_data$employment_area == "it_and_information_services", "IT", 
                             ifelse(summary_data$employment_area == "Missing", "Missing Data", summary_data$employment_area))
summary_data <- summary_data %>%
    mutate(percentage = Count / sum(Count) * 100)

# Determine the maximum count value for outline (which corresponds to "IT" or "Missing Data")
max_value_row <- which.max(summary_data$Count)

# Plot the pie chart
ggplot(summary_data, aes(x = "", y = Count, fill = label)) +
    geom_bar(stat = "identity", width = 1, color = ifelse(1:nrow(summary_data) == max_value_row, "black", NA), size = 1.2) +  # Add outline for max value
    coord_polar(theta = "y") +
    geom_text(aes(label = ifelse(1:nrow(summary_data) == max_value_row, 
                                 paste0(label, "\n", round(percentage), "%"), 
                                 "")), 
              position = position_stack(vjust = 0.5), size = 5, color = "white") +
    theme_minimal() +
    labs(title = "Learners per Employment Area", fill = "Employment Area") +
    theme(
        axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        plot.title = element_text(hjust = 0.5)
    ) +
    scale_fill_manual(values = colors) +  # Apply the custom color palette
    guides(fill = guide_legend(title.position = "top", title.hjust = 0.5, label.position = "right"))
```

The remaining 91% (roughly 57,556 learners) do not have employment area information. Because of which we cannot come up with which area of employers take up this cyber security course. 

There are around 63,248 learners with respect to the Table 1, who enrolled altogether in the cyber security course. But the unique number of learners are only 35,225 according to Table 1. Which means the rest 28,023 are either duplicate or re-joiners. 

So to investigate on this I am trying to plot a visualization where I have grouped the re-joiners 2-4 as one bucket and 5-8 rejoining count learners as another bucket. And from the visualization below we can conclude that there are more number of re-joiners who are coming under the bucket count 2-4. 

```{r,message=FALSE, warning=FALSE, echo = FALSE,fig.height=4,fig.width=5,fig.align='center'}
duplicate_counts <- duplicate_counts %>%
  mutate(rejoin_bucket = case_when(
    count >= 2 & count <= 4 ~ "2-4 Rejoins",
    count >= 5 & count <= 8 ~ "5-8 Rejoins",
    count > 8 ~ "9+ Rejoins",
    TRUE ~ "1 Rejoin"
  ))

# Count the number of learners in each bucket
bucket_counts <- duplicate_counts %>%
  group_by(rejoin_bucket) %>%
  summarise(learner_count = n())

# Plotting the bucket counts
ggplot(bucket_counts, aes(x = rejoin_bucket, y = learner_count, fill = rejoin_bucket)) +
  geom_bar(stat = "identity") +
  scale_fill_brewer(palette = "Set3") +  # Using a user-friendly color palette
  labs(
    title = "Learners Rejoining the Course",
    x = NULL,
    y = NULL
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_blank(),  # Remove x-axis labels
    axis.ticks.x = element_blank(),  # Remove x-axis ticks
    plot.title = element_text(hjust = 0.5)
  )
```

Since there are many number of re-joiners for the cyber security course. Now, looking at the survey response collected from the learners for leaving the course I have tabled number of learners reasons to leave the course with unique reasons and its total count chosen by the learners.

```{r,message=FALSE, warning=FALSE, echo = FALSE}
unique_leaving_reasons_count <- duplicates_data_leaving_survey_responses %>%
  count(leaving_reason) %>%    # Count occurrences of each leaving reason
  arrange(leaving_reason)      # Optionally, arrange alphabetically

# Display the unique leaving reasons with their counts in a table format using kable
kable(unique_leaving_reasons_count, col.names = c("Unique Leaving Reason", "Count"))
```

From the above table it is stated that most of the learners feel that the course is very lengthy that they do not find enough time to complete. Which is why they are re-joining it or maybe the course is framed perfect its the learners are not able to frame time for learning. Nearly 144 learners selected 'I prefer not to say' and 'Others' and 40 of them selected 'The course required more time than I realized' which altogether sums up 359 of them feel like time is the problem. Only 72 learners expectations did not match with the course and 26 of them found it hard.
